Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-16 doc/GMP2015/s01_introduction.tex
Slight changes to wording. Added some references.

Some properties of natural and engineered materials can be analyzed by
treating them as stacked two dimensional (2D) layers. The structure
within each layer could be self-similar \cite{Intro1} spanning
multiple scales, generally aperiodic and quasi-uniform within any one
scale, and consisting of a few repeated motifs appearing in disordered
arrangements. Each layer is not necessarily planar, i.e., it consists
of multiple, inter-constraining planar (genus 0) monolayers.
Furthermore, a layer is often  either {\sl isostatic or
underconstrained, not self-stressed} (See Section \ref{sec:prelim} for
definitions). These properties are consistent with a self-assembled 2D
structure that minimizes mass, optimally distributes external stresses
and itself participates in the assembly of diverse and
multifunctional, larger 2D structures.

@@

    \item crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}, as well as crosslinked actin filaments in the cytoskeleton matrix.

    \item more recent, engineered examples including  disordered graphene layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced by  microfibrils; and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet, bringing other self-similar structures \cite{Microfibrils2} within reach.

    \item  silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 2D particles under non-overlap constraints, e.g, jammed disks on the plane \cite{JammedDisk1}
\end{enumerate}
%
In order to study structural and mechanical properties such as
rigidity, flexes and distribution of external stresses within the
material layer, it is natural to model a material layer as a solution
of a geometric constraint system of appropriate types of geometric
primitives, under metric or algebraic constraints. Such  2D {\bf{\em
qusecs}} (quasi-uniform or self-similar constraint systems) can be
used to understand or design material layers (their solutions) with
desired properties.
%
\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
We now briefly survey existing techniques for studying 2D {\em
qusecs,}  many of which are {\it bar-joint} systems (Examples 1,2
above, see Sections \ref{sec:DRP}, \ref{sec:recomb}), {\it multibody-
pin} systems (Example 4,5, see Section \ref{sec:bodypin}) or  {\it
pinned-line incidence} systems (Example 3, see Section
\ref{sec:pinnedline}). The limitations of these techniques directly
motivate the contributions of this paper.

\meeraheader{i}{Finding (vertex)-maximal generically rigid subsystems}
Fast, graph-based algorithms exist (pebble-game \cite{XX}), for
locating all maximal, generically rigid subsystems \cite{XX}. When the
input is rigid, these algorithms compute the identity and  output the
input.

@@

\meeraheader{ii}{Optimal Recursive Decomposition (DR-planning)}
Recursive decomposition of geometric constraint systems has been
formalized \cite{XX} and well-studied \cite{XX} as the {\sl
Decomposition-Recombination (DR)-planning} problem. For the above-
mentioned classes of 2D qusecs, generic rigidity is a combinatorial
property and hence each level of the decomposition should, in
principle, be achievable by a graph-based algorithm as in (1), without
involving geometric information in the constraint system. Since many
such decompositions can exist for a given constraint system, criteria
defining desirable or optimal DR-plans and DR-planning algorithms were
given in \cite{XX}. An {\em optimal DR-plan} is one that minimizes the
maximum number of child subsystems of any parent system, a property
that is crucial for obtaining a solution to the parent constraint
system from the solutions of the child systems.

@@

\meeraheader{iii}{DR-plans for special classes and with other
criteria}
For a special class of 2D qusecs, namely  {\em tree-decomposable}
systems \cite{XX} common in computer aided mechanical design, (which
includes ruler-and-compass and Henneberg-I constructible systems), all
DR-plans turn out to be optimal. This satisfies the so-called {\em
Church-Rosser} criterion, leading to highly efficient DR-planning
algorithms. For general 2D qusecs, alternate criteria were suggested
such as {\em cluster minimality} requiring parent systems to be
composed of a minimal set of at least 2 rigid proper subsystems (i.e.,
no proper subset forms a rigid system); and {\em proper maximality},
requiring child subsystems to be maximal rigid proper subsystems of
the parent system.

@@

However, the generality of these algorithms trades-off against
efficiency, and despite the optimization, the best algorithms can
still take exponential time in the number of child subsystems (which
can be arbitrarily large even for optimal DR-plans) in order to
guarantee all solutions of a given orientation type, even for a single
(sub)system in a DR-plan. They are prohibitively slow in practice. We
note that utilizing the DR-plan and optimal recombination as a
principled basis, high performance heuristics and software exists
\cite{XX,XX} to tame combinatorial explosion via user intervention.

\meeraheader{v}{Configuration Spaces of Underconstrained Systems}
For underconstrained 2D bar-joint and multibody-pin cusecs obtained
from various subclasses of tree-decomposable systems, algorithms have
been developed \cite{XX, XX} that to complete them into isostatic
systems \cite{XX, XX} and to find paths within the connected
components \cite{XX,XX} of standard cartesian configuration spaces.
Most of the algorithms with formal guarantees leverage Cayley
configuration space theory \cite{XX,XX} to characterize subclasses of
graphs and additional constraints that control combinatorial
explosion, and provide faithful bijective representation of connected
components and paths. These algorithms have decreasing efficiency as
the subclass of systems gets bigger,  with highest efficiency for
underlying partial 2-tree graphs (alternately, tree-width 2, series-
parallel, $K_4$ minor avoiding), moderate efficiency for 1 degree-
of-freedom (dof) graphs with low Cayley complexity (which include
common linkages such as the Strandbeest, Limacon and Cardioid), and
decreased efficiency for general 1-dof tree-decomposable graphs. While
software suites exist \cite{XX,XX}, no such formal algorithms and
guarantees are known for non-tree-decomposable systems.
%
\subsection{Contributions and Organization}
\label{sec:cont}

The contributions of this paper are the following.
\begin{itemize}
\item
We navigate the NP-hardness barrier mentioned in (2) above, for
finding optimal DR-plans by defining a so-called {\it canonical} DR-
plan and showing a strong Church-Rosser property: {\it all canonical
DR-plans for isostatic or underconstrained 2D qusecs are optimal}
(Section \ref{sec:DRP}).

\item
We give an efficient ($O(n^2)$) algorithm to find a canonical (and
hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above
(Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}.
The canonical DR-plan illuminates the essence of the NP-hardness of
finding optimal DR-plans for overconstrained systems.

\item
We give a method to deal with the algebraic complexity of recombining
the realizations or solutions of  child subsystems into a solution of
the parent system. Specifically, we define the problem of  minimally
modifying the indecomposable recombination system so that it becomes
decomposable via a small DR-plan and yet preserves the original
solutions in an efficiently searchable manner. We show formal
connection to well known problems such as optimal completion of
underconstrained  systems  \cite{XX}.  When the required modifications
are bounded, we obtain new, efficient algorithms for realizing both
isostatic and underconstrained qusecs by leveraging recent results
about Cayley parameters in \cite{XX,XX}  (Section \ref{sec:recomb}).

\item
We briefly describe applications of the above techniques to modeling,
analyzing and designing specific properties in 2D material layers. For
Examples 1,2 (achieving isostaticity, distribution of stresses in
self-similar, bar-joint systems); Example 3 (canonical and optimal DR-
plans for pinned line incidence systems \cite{XX}) and Examples 4,5
(boundary-conditions for achieving various desired properties of
multi-body  pin systems).

>>>>

Some properties of natural and engineered materials can be analyzed by
treating them as stacked two dimensional (2D) layers. The structure
within each layer could be self-similar \cite{Intro1} spanning
multiple scales, generally aperiodic and quasi-uniform within any one
scale, and consisting of a few repeated motifs appearing in disordered
arrangements. Each layer is not necessarily planar, i.e., it consists
of multiple, inter-constraining planar (genus 0) monolayers.
Furthermore, a layer is often  either {\sl isostatic or
under-constrained, not self-stressed} (See Section \ref{sec:prelim} for
definitions). These properties are consistent with a self-assembled 2D
structure that minimizes mass, optimally distributes external stresses
and itself participates in the assembly of diverse and
multifunctional, larger 2D structures.

@@

    \item Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls \cite{CellWalls1} \cite{CellWalls1}, as well as crosslinked actin filaments in the cytoskeleton matrix.

    \item More recent, engineered examples, including disordered graphene layers \cite{Graphene1} \cite{Graphene2} sometimes reinforced by  microfibrils; and DNA assemblies \cite{Microfibrils1} including a recent Sierpinski carpet, bringing other self-similar structures \cite{Microfibrils2} within reach.

    \item Silica bi-layers, glass \cite{SilicaGlass1} \cite{SilicaGlass2}, and materials that behave like assemblies of 2D particles under non-overlap constraints, e.g, jammed disks on the plane \cite{JammedDisk1}
\end{enumerate}
%
In order to study structural and mechanical properties such as
rigidity, flexes and distribution of external stresses within the
material layer, it is natural to model a material layer as a solution
of a geometric constraint system of appropriate types of geometric
primitives, under metric or algebraic constraints. Such 2D {\bf{\em
qusecs}} (quasi-uniform or self-similar constraint systems) can be
used to understand or design material layers (their solutions) with
desired properties.

\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
We now briefly survey existing techniques for studying 2D
\textit{qusecs},  many of which are \textit{bar-joint} systems
(Examples 1, 2 above, see Sections \ref{sec:DRP}, \ref{sec:recomb}),
\textit{body-hyperpin} systems (Example 4,5, see Section
\ref{sec:bodypin}) or \textit{pinned-line incidence} systems (Example
3, see Section \ref{sec:pinnedline}). The limitations of these
techniques directly motivate the contributions of this paper.

\meeraheader{i}{Finding (vertex)-maximal generically rigid subsystems}
Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG}), for
locating all maximal, generically rigid subsystems \cite{XX}. When the
input is rigid, these algorithms compute the identity and output the
input.

@@

\meeraheader{ii}{Optimal Recursive Decomposition (DR-planning)}
Recursive decomposition of geometric constraint systems has been
formalized \cite{XX} and well-studied \cite{XX} as the {\sl
Decomposition-Recombination (DR)-planning} problem. For the above-
mentioned classes of 2D qusecs, generic rigidity is a combinatorial
property and hence each level of the decomposition should, in
principle, be achievable by a graph-based algorithm as in (1), without
involving geometric information in the constraint system. Since many
such decompositions can exist for a given constraint system, criteria
defining desirable or optimal DR-plans and DR-planning algorithms were
given in \cite{XX}. An \emph{optimal DR-plan} is one that minimizes
the maximum number of child subsystems of any parent system, a
property that is crucial for obtaining a solution to the parent
constraint system from the solutions of the child systems.

@@

\meeraheader{iii}{DR-plans for special classes and with other
criteria}
For a special class of 2D qusecs, namely \emph{tree- decomposable}
systems \cite{XX} common in computer aided mechanical design, (which
includes ruler-and-compass and Henneberg-I constructible systems), all
DR-plans turn out to be optimal. This satisfies the so-called \emph
{Church-Rosser} criterion, leading to highly efficient DR-planning
algorithms. For general 2D qusecs, alternate criteria were suggested
such as \emph{cluster minimality} requiring parent systems to be
composed of a minimal set of at least 2 rigid proper subsystems (i.e.,
no proper subset forms a rigid system); and \emph{proper maximality},
requiring child subsystems to be maximal rigid proper subsystems of
the parent system.

@@

However, the generality of these algorithms trades-off against
efficiency, and despite the optimization, the best algorithms can
still take exponential time in the number of child subsystems (which
can be arbitrarily large even for optimal DR-plans) in order to
guarantee all solutions of a given orientation type, even for a single
(sub)system in a DR-plan. They are prohibitively slow in practice. We
note that, utilizing the DR-plan and optimal recombination as a
principled basis, high performance heuristics and software exists
\cite{XX,XX} to tame combinatorial explosion via user intervention.

\meeraheader{v}{Configuration Spaces of Under-constrained Systems}
For under-constrained 2D bar-joint and body-hyperpin qusecs obtained
from various subclasses of tree-decomposable systems, algorithms have
been developed \cite{XX, XX} to complete them into isostatic systems
\cite{XX, XX} and to find paths within the connected components
\cite{XX,XX} of standard Cartesian configuration spaces. Most of the
algorithms with formal guarantees leverage Cayley configuration space
theory \cite{XX,XX} to characterize subclasses of graphs and
additional constraints that control combinatorial explosion, and
provide faithful bijective representation of connected components and
paths. These algorithms have decreasing efficiency as the subclass of
systems gets bigger, with highest efficiency for underlying partial
2-tree graphs (alternately, tree-width 2, series- parallel, $K_4$
minor avoiding), moderate efficiency for 1 degree- of-freedom (dof)
graphs with low Cayley complexity (which include common linkages such
as the Strandbeest, Limacon and Cardioid), and decreased efficiency
for general 1-dof tree-decomposable graphs. While software suites
exist \cite{XX,XX}, no such formal algorithms and guarantees are known
for non-tree-decomposable systems.

\subsection{Contributions and Organization}
\label{sec:cont}

The contributions of this paper are the following.
\begin{itemize}
  \item We navigate the NP-hardness barrier mentioned in (2) above,
  for finding optimal DR-plans by defining a so-called \textit{canonical}
  DR-plan and showing a strong Church-Rosser property: \textit{all
  canonical DR-plans for isostatic or under-constrained 2D qusecs are
  optimal} (Section \ref{sec:DRP}).

  \item We give an efficient ($O(n^2)$) algorithm to find a canonical
  (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned
  above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and
  \ref{sec:pinnedline}). The canonical DR-plan illuminates the essence
  of the NP-hardness of finding optimal DR-plans for over-constrained
  systems.

  \item We give a method to deal with the algebraic complexity of
  recombining the realizations or solutions of child subsystems into a
  solution of the parent system. Specifically, we define the problem
  of minimally modifying the indecomposable recombination system so
  that it becomes decomposable via a small DR-plan and yet preserves
  the original solutions in an efficiently searchable manner. We show
  formal connection to well known problems such as optimal completion
  of under-constrained systems \cite{XX}. When the required
  modifications are bounded, we obtain new, efficient algorithms for
  realizing both isostatic and under-constrained qusecs by leveraging
  recent results about Cayley parameters in \cite{XX,XX} (Section
  \ref{sec:recomb}).

  \item We briefly describe applications of the above techniques to
  modeling, analyzing, and designing specific properties in 2D material
  layers. For Examples 1,2 (achieving isostaticity, distribution of
  stresses in self-similar, bar-joint systems); Example 3 (canonical
  and optimal DR-plans for pinned line incidence systems \cite{XX})
  and Examples 4,5 (boundary-conditions for achieving various desired
  properties of body-hyperpin systems).
